---
title: "Homework 3"
author: "Prajwal Nagaraju"
date: "January 29, 2017"
output: pdf_document
---

#Set Working directory ,read data
```{r}
getwd()
setwd("~/Desktop")
bike <- read.csv(file = "day.csv")
str(bike)
library(caret)
library(arm)
library(dplyr)
library(ggplot2)
```
#With cnt as the outcome variable, fit a K-nearest neighbor regression model (KNN) using the train()
#function in the caret package with the following variables as predictors: mnth, holiday, weathersit, temp,
#and windspeed. Report RMSE and R-squared. Round your results to two decimals.
```{r}
#KNN model
(knn_model <- train(cnt ~ mnth + holiday + weathersit + temp + windspeed , 
      preProcess = c("center","scale"),  #have to center and scale for KNN
      method="knn",
      data =bike))

#Compute RMSE
rmse <- function(yhat, y) {
  sqrt((mean((yhat - y)^2)))
}

knn_rmse <- round(rmse(bike$cnt, predict(knn_model)) ,2)
knn_rmse

#Compute R2
R2 <- function(y, yhat, ybar, digits = 2) {
  round(1 - sum((y - yhat)^2)/sum((y - ybar)^2), digits)
}

knn_R2 <- round(R2(y = bike$cnt, yhat = predict(knn_model), mean(bike$cnt)) ,2)
knn_R2
```

#Using the same variables as above, fit a linear model. Note that the categorical variables with more than
#two levels need to be recoded, either as character or factor variables, otherwise R will misunderstand
#them as numeric. lm() will treat character variables appropriately as factors. Report RMSE and
#R-squared for this model. Round your results to two decimals.
```{r}
#convert to factor
bike$mnth_factor <- factor(bike$mnth)
#b$holiday <- factor(b$holiday)
bike$weathersit_factor <- factor(bike$weathersit)

#LM model
(lm_model <- train(cnt ~ mnth_factor + holiday + weathersit_factor + temp + windspeed, 
                   method="lm",
                   data =bike))

#round(summary(lm_model)$r.sq ,2)
#round(summary(lm_model)$sigma ,2)

#Compute RMSE
lm_rmse <- round(rmse(y = bike$cnt,yhat = predict(lm_model)) ,2)
lm_rmse

#Compute R2
lm_R2 <- round(R2(y = bike$cnt, yhat = predict(lm_model), mean(bike$cnt)) ,2)
lm_R2
```

#In class we have noted that the temp variable is not quite linear when plotted against cnt. Mean
#ridership increases as temperature increases, then levels off, then declines at the highest temperatures.
#Create a new variable, temp2, by multiplying temp x temp. Refit the KNN and linear models above
#using this additional variable. Report RMSE and R2 for both models. Round your results to two
#decimals.

```{r}
#Create a new variable, temp2, by multiplying temp x temp.
bike$temp2 <- (bike$temp) * (bike$temp)

#KNN Model
(knn_model2 <- train(cnt ~ mnth + holiday + weathersit + temp + temp2 + windspeed , 
      preProcess = c("center","scale"),  #have to center and scale for KNN
      method="knn",
      data =bike))

#Compute RMSE
knn_model2_RMSE <- round(rmse(bike$cnt, predict(knn_model2)) ,2)
knn_model2_RMSE

#Compute R2
knn_model2_R2 <- round(R2(y = bike$cnt, yhat = predict(knn_model2), mean(bike$cnt)),2)
knn_model2_R2

#LM model

(lm_model2 <- train(cnt ~ mnth_factor + holiday + weathersit_factor + temp + temp2 + windspeed, 
                   method="lm",
                   data =bike))

#Compute RMSE
lm_model2_RMSE <- round(rmse(bike$cnt, predict(lm_model2)) ,2)
lm_model2_RMSE

#Compute R2
lm_model2_R2 <- R2(y = bike$cnt, yhat = predict(lm_model2), mean(bike$cnt))
lm_model2_R2

#round(summary(lm_model2)$r.sq ,2)
#round(summary(lm_model2)$sigma ,2)
```

#Using the linear model you fit above (with mnth, holiday, weathersit, temp and windspeed as explanatory
#variables, but not temp2), predict ridership for a holiday in May when temperature and windspeed are
#average and weathersit = 1. Round your results to the nearest whole number.

```{r}
(fit <- lm(cnt ~ mnth_factor + holiday + weathersit_factor + temp + windspeed, data = bike))
display(fit)
#predict ridership for a holiday in May when temperature and windspeed are average and weathersit = 1.
value <- round(1694.59 + (961.41 * 1) + (-769.06 * 1) + (5867.27 * mean(bike$temp)) + (-2690.61 * mean(bike$windspeed)))
value
```

#Using the same model (i.e., without the temp2 variable), calculate 95% confidence intervals for the
#coefficient for temp. Round your results to two decimals.

```{r}
fit <- lm(cnt ~ mnth_factor + holiday + weathersit_factor + temp + windspeed, data = bike) 
display(fit)

#95% confidence intervals for the coefficient for temp

CI1 <- round(5867.27 + 1.96 * 669.56 ,2)
CI2 <- round(5867.27 - 1.96 * 669.56 ,2)

CI1
CI2
```

#Fit two simple bivariate models of cnt using temp as the predictor. For the first, use KNN
#regression, and for the second use linear regression. Fit a third regression using linear regression but add
#a quadratic term for temp. Upload three scatter plots that display ^??(t) as estimated non-parametrically
#by the KNN algorithm and parametrically by the ordinary least squares algorithm, with and without
#the quadratic term for temp. In the plots, temp should be on the x-axis and cnt on the y-axis. Show
#^??(t) with a red line. Title and label your plots appropriately.
```{r}
# FIt a KNN model

(fit_knn_cnt <- train(cnt ~ temp ,
      preProcess = c("center","scale"),  #have to center and scale for KNN
      method="knn",
      data =bike))

#Fit a Linear Model 
(fit_lm_cnt <- train(cnt ~ temp, 
                   method="lm",
                   data =bike))

#Scatterplot:
plot(bike$temp, bike$cnt , main = "Count as a function of Temperature" ,xlab = "Temperature" ,ylab ="Ridership")
abline(lm(cnt ~ temp, data = bike), col = 2)

#Fir a linear model with quadratic equation
(fit_lm_cnt2 <- train(cnt ~ temp2 + temp, 
                   method="lm",
                   data =bike))

#Scatterplot:
plot(bike$temp, bike$cnt , main = "Count as a function of Temperature" ,xlab = "Temperature" ,ylab ="Ridership")
abline(gam(cnt ~ temp + temp2, data = bike), col = 2)

```

