---
title: "Homework1_PN"
author: "Prajwal Nagaraju"
date: "January 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read data
```{r}
bikeday <- read.csv(file = "day.csv")
```

#Create and upload a histogram of bike ridership (encoded in the variable "cnt) for 2011.
```{r}
hist(subset(bikeday,yr == 0)$cnt,xlab = "Count" ,ylab = "Values" ,main="bike ridership for 2011")
box()
```

#Create and upload a histogram of bike ridership (encoded in the variable "cnt) for 2012.
```{r}
hist(subset(bikeday,yr == 1)$cnt,xlab = "Count" ,ylab = "Values" ,main="bike ridership for 2012")
box()
```

#Create and upload a boxplot of "cnt" by year.
```{r}
boxplot(bikeday$cnt ~ bikeday$yr ,main = "boxplot by year" , xlab = "Year" ,ylab = "Count")
```

#Create and upload a boxplot of "cnt" by workingday.
```{r}
boxplot(bikeday$cnt ~ bikeday$workingday ,main = "boxplot by working day" ,xlab = "Working day" ,ylab = "Count")
```

#Create and upload a boxplot of "cnt" by month.
```{r}
boxplot(bikeday$cnt ~ bikeday$mnth ,main = "boxplot by month" ,xlab ="Month" ,ylab= "count")
```

#Which month had the highest median number of riders?
```{r}
a <- c(1)
for(i in 1:12){
a[i] <- c(median(subset(bikeday$cnt,bikeday$mnth == i)))
}
order(a ,decreasing = T)


bikeday %>%
group_by(mnth) %>%
summarize(med = median(cnt)) %>%
filter(med==max(med)) %>%
kable()
```

#Which month, based on the middle 50% of observations, had the most riders?
```{r}
b <- c(1)
for(i in 1:12){
b[i] <- c(IQR(subset(bikeday$cnt ,bikeday$mnth == i)))
}
order(b ,decreasing = T)
```


#Turn yr into a factor variable and create and upload a scatterplot with temp on the x-axis and cnt on
#the y-axis. Color the points by yr. Challenge: create legend for color.
```{r}
bikeday$yr <- factor(bikeday$yr)
str(bikeday)
plot(bikeday$temp ,bikeday$cnt ,xlab = "Temperature" ,ylab = "Count" ,main ="Relationship between temp and count by year" ,col= bikeday$yr)

#legend
year <- bikeday$yr
levels(year) <- c("2011" ,"2012")
legend('topleft', legend = levels(year),  col = 1:2, cex = 1, pch = 1)
```

#Create two histograms of cnt, one for winter and one for summer, and put them together, one on top of
#the other. Upload that single plot. Make sure that the x axes have the same scale.
```{r}
hist(subset(bikeday ,season == 2)$cnt,xlab = "Count" ,ylab = "Values" ,main = "Overlapping Histogram for Summer and Winter")
hist(subset(bikeday ,season == 4)$cnt ,add = T)
box()
```


