---
title: "Homework2"
author: "Prajwal Nagaraju"
date: "January 25, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read Data
```{r}
bike <- read.csv(file = "day.csv")
library(dplyr)
library(ggplot2)
```

#Create and upload a histogram of bike usage that is faceted by temperature, which you have turned
#into a binary variable: above average (or equal to the average) and below average. Make sure that
#"below average" is in the left column and "above average" is in the right. The way to do this is to
#re-order the default (alphabetic) factor levels of your binary temperature variable.
```{r}
bike %>% 
  mutate(temp_bin = ifelse(temp <= mean(temp), "below_avg_temp", "above_avg_temp")) %>%
    ggplot(aes(cnt)) + geom_histogram() + facet_wrap(~factor(temp_bin ,levels = c("below_avg_temp" ,"above_avg_temp"))) +
      ggtitle("histogram of Ridership by temperature") + labs(x = "Ridership" ,y = "Values")
```

#Create and upload a histogram of bike usage that is faceted by month and the binary temperature
#variable: above average (or equal to the average) and below average. Month should be organized as
#rows, temperature as columns. Again make sure that "below average" is in the left column and "above
#average" is in the right.
```{r}
bike %>% mutate(temp_bin = ifelse(temp <= mean(temp), "below_avg_temp", "above_avg_temp")) %>%
  ggplot(aes(cnt)) + geom_histogram() + facet_grid(mnth~factor(temp_bin ,levels = c("below_avg_temp" ,"above_avg_temp"))) +
    ggtitle("histogram of Ridership by temperature and Month") + labs(x = "Ridership" ,y = "Values")
```

#Create and upload a similar plot-a histogram of bike usage that is faceted by month and a binary
#temperature variable. This time, however, make sure that the binary temperature variable, above and
#below average, is computed uniquely for each month. For example, the values of the binary temperature
#variable for January should be above and below average for January. Again, make sure that "below
#average" is in the left column and "above average" is in the right.
```{r}
bike %>%
  group_by(mnth) %>%
    mutate(temp_bin = ifelse(temp <= mean(temp), "below_avg_temp", "above_avg_temp")) %>%
      ggplot(aes(cnt)) + geom_histogram() + facet_grid(mnth~factor(temp_bin ,levels = c("below_avg_temp" ,"above_avg_temp"))) +
        ggtitle("histogram of Ridership by temperature grouped by Month") + labs(x = "Ridership" ,y = "Values")
```

#What are the 95% confidence intervals for the conditional mean of ridership in each season? Round
#your answers to whole numbers.
```{r}
bike %>% 
  group_by(season) %>%
    summarise(mean_season = mean(cnt) ,sem_season = sd(cnt)/ sqrt(length(cnt)) ,CI1 = round(mean_season - sem_season * 1.96) ,CI2 = round(mean_season + sem_season * 1.96)) %>%
      select(season ,CI1 ,CI2)
```

#Re-create the above plot exactly using your own dplyr and ggplot code.
```{r}
s <- bike %>%
  group_by(mnth) %>%
  mutate(monthly_temp = as.factor(ifelse(temp >= mean(temp), "Above average", "Below average")))

s %>%
  group_by(mnth, monthly_temp)%>%
  summarise(mean_cnt = mean(cnt))%>%
  ggplot(aes(x = factor(mnth), y = mean_cnt, group = monthly_temp, col = factor(monthly_temp, levels(monthly_temp)[c(2,1)]))) + scale_color_discrete(name = "monthly temp") + geom_line() +xlab("month") + ylab("ridership") +
  ggtitle("Average Ridership by month and temperature")
```

